Benchmarking on Rigetti architecture

### 2022-08-02 txt/manuscript.tex
manuscript changes - proofs

\subsubsection{Model specification}
Theoretically, scenario of this set up assume that Alice and Bob have an 
unknown measurement device, black--box.  Alice has hiden one of the two 
measurements  in the black box.  The only information which Bob has is that it 
performs one of such  measurements. 
The  goal of discrimination of von Neumann measurements  is to determine by Bob 
whether it is possible to discriminate measurements perfectly, i.e. with 
probability equals one. If this is not possible, he  would like to know the 
upper bound of such a probability.  Theoretical results of discrimination of 
von Neumann measurements is well-known and characterized 
in~\cite{puchala2018strategies}.    The scheme of von Neumann discrimination is 
presented in Fig.~\ref{fig:scheme} and 
can be divided into the following steps:

\begin{enumerate}
	\item Alice and Bob share entanglement state $\ket{\psi_{AB}}$.
	\item Alice performs one of two known measurements on part $A$ of the input 
	state  $\ket{\psi_{AB}}$.
	\item Alice  uses
	the output label $i$ and performs a conditional binary measurement on part 
	$B$.
	\item  Bob measures the part $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision  basis of the condition binary measurement.
\end{enumerate}

\begin{figure}[h]
	\centering

	\includegraphics[width=0.9\linewidth]{gates1.pdf} 
	\caption{A schematic representation of the setup for distinguishing
		measurements by using entanglement. }
	\label{fig:scheme}
\end{figure}

\subsubsection{Rigetti architecture's limits/ practical solutions}

We may encounter numerous obstacles in the implementation process due to the 
limited architecture of quantum computer. 
Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:diamond}). One of the components of such a 
scheme is to perform the conditional binary measurement. The first limitation 
is lack to possibility of implementation of the conditional binary measurement. 
The only possible measurement performed on quantum architectures is measurement 
in the computational basis. 
A postselection method provides us with a clever going around the obstacle. The 
postselection  scheme (see Fig.~\ref{postselection})  can be described as 
follow.

\begin{enumerate}
	\item We prepare input state $\ket{\psi_{AB}}$.
	\item One of two measurements is performed on part A.
	\item We prepare optimal measurements $V_j$, which are well known by 
	Holevo--Helstrom theorem and perform them on part $B$.
	\item We get the output label $i$.
	\item We include these cases for which $i = j$. Otherwise, we reject them.
	\item By the use of $V_j$ output, we decide whether the measurement was 
	performed on part $A$.
	\item We count the probability of postselection.
\end{enumerate}

\todo[inline]{PAMIETAJ O DAGGERACH !!!}

\begin{figure}[h!]
	\includegraphics[scale=1.5]{onequbit.pdf}
	\caption{A schematic representation of the setup for distinguishing
		measurements using postselection.}
	\label{postselection}
\end{figure}

The schematic representation of theoretical setup is shown in the following 
figure.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.2]{pics/theoretical_scheme}
	\caption{Theoretical discrimination scheme}
	\label{fig:theoretical_scheme}
\end{figure}
\todo[inline]{konretne przyklady zobaczymy potem... }

\subsubsection{Two schemes of practical realization}

\begin{scheme}(By using postsellection)

@@

		\caption{ A schematic representation of the setup for distinguishing
			measurements using postselection. 
		}\label{fig:postsellection}
	\end{figure}
	\begin{enumerate}
		\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
		\rangle. $
		\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
		\item We implement unitary $V_0 $ or $ V_1$.
		\item We prepare the measurement $\Delta$ in standard basis (already exists on 
		Rigetti architecture) on each qubits.
		\item We calculate the probability of correct discrimination in the following 
		way
		\begin{equation}
		p = ...
		\end{equation}
		%\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}
	\end{enumerate}
\end{scheme}

@@

\subsubsection{Theoretical results}

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity notation, square matrices will be 
denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$ 
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite 
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is 
positive semidefinite operators having
trace equal to one, will be denoted $\Omega_d$.  The subset of $M_d$ consisting of 
unitary matrices will be denoted by
$\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$.  We will also need a linear
mapping transforming $M_{d_1}$ into $M_{d_2}$, which will be denoted \begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation} 
\todo[inline]{nie mowimy channel} 
Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
Mostly, we will use the unitary channels
$\Phi_{U}$ which are defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
expressed as the action of a quantum channel
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$ Every von Neumann measurement  $\PP_{U}$ can be also rewritten as $\Delta 
\circ \Phi_{(UE)^\dagger}$, where $\Delta$ denotes the completely dephasing 
channel.

\subsubsection{Diamond norm}
Let us now consider linear mappings transforming  square matrices into square 
matrices 
that is \ $\Phi: M_{d_1} \to M_{d_2}$. 
 We define its completely bounded trace
norm, also known as a diamond norm, as
\begin{equation}
\|\Phi\|_\diamond = \max_{\|X\|_1=1} \| \left(\Phi \otimes \1\right) (X) \|_1.
\end{equation}

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum operations, $\TT$  and $\SS$, 
in terms of their distance with the use of the diamond norm
\begin{equation}
p =  \frac12 + \frac14 \| \TT - \SS \|_\diamond.
\end{equation}

>>>>

\subsubsection{Model specification}
Theoretically, the scenario of this set-up assumes that Alice and Bob have an 
unknown measurement device, a black--box. Alice has hidden one of the two 
measurements, say $\SS$ and $\TT$,  in the black box. Bob only has information that it 
performs one of such measurements. 
First, they want to tell whether it is possible to discriminate $\SS$ and $\TT$ perfectly, i.e. with probability equal to one. If this is not the case, we would like to
know the upper bound of such a probability. 
Theoretical results of discrimination of 
von Neumann measurements are well-known and characterized 
in~\cite{puchala2018strategies}.   
They devised the optimal discrimination strategy between two von Neumann measurements and also calculated the exact probability of their discrimination.

In this work we will consider the problem of discrimination between the von Neumann measurement in the computational basis  and  the von Neumann measurement in basis $U$, where $U$ is a unitary matrix. 
This scheme is 
presented in Fig.~\ref{fig:scheme} and 
can be divided into the following steps:

\begin{figure}[h!]
	\centering 
	\includegraphics[width=0.9\linewidth]{gates1.pdf} 
	\caption{A schematic representation of the setup for distinguishing
		measurements by using entanglement. }
	\label{fig:scheme}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.5]{pics/theoretical_scheme}
	\caption{Theoretical discrimination scheme}
	\label{fig:theoretical_scheme}
\end{figure}
\todo[inline]{tan powinien byc $\PP_{V_i}$}

\begin{enumerate}\item Alice and Bob share entanglement state $\ket{\psi_{AB}}$.
	\item Alice performs one of two known measurements,  say $\PP_{\Id}$ and $\PP_{U}$,  on part $A$ of the input 
	state  $\ket{\psi_{AB}}$.
	\item Alice  uses
	the output label $i$ and performs the conditional binary measurement on the part 
	$B$.
	\item  Bob measures the part $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on received label $j$. If $j=0$, he decides that the measurement $\PP_U$ is hidden in the black box. Otherwise, he decides that the measurement $\PP_{\Id}$ was hidden in the black box. 
\end{enumerate}

\subsubsection{Quantum architecture's limits and practical solutions}

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:scheme}). One of the components of such a 
scheme is to perform the conditional binary measurement. The first limitation 
is a lack of possibility of implementing the conditional binary measurement. 
The only possible measurement performed on quantum architectures is the measurement 
on a computational basis. 
Here, we present two possible methods that provide a clever way to cross the obstacle (two schemes of practical realization).

The first way is a postselection scheme which can be described as 
follow.  The schematic representation of this setup we can see on Fig.~\ref{postselection}.     
\begin{scheme}(By using postsellection)

\todo[inline]{PAMIETAJ O DAGGERACH  no i u nas to bedzie nazwane $\PP_{V^\dagger_j}$!!!} 
\begin{figure}[h!]
	\includegraphics[scale=1.5]{onequbit.pdf}
	\caption{A schematic representation of the setup for distinguishing
		measurements using postselection.}
	\label{postselection}
\end{figure}

@@

	\caption{ A schematic representation of the setup for distinguishing
		measurements using postselection. 
	}\label{fig:postsellection}
\end{figure} 
\begin{enumerate}
	\item We prepare input state $\ket{\psi_{AB}}$.
	\item One of two measurements is performed on part A.
	\item We prepare optimal measurements $V_j$, which are well known by 
	Holevo--Helstrom theorem and perform them on part $B$.
	\item We receive an output label $i$.
	\item We include these cases for which $i = j$. Otherwise, we reject them.
	\item By the use of $V_j$ output, we decide whether the measurement was 
	performed on part $A$.
	\item We count the probability of postselection.
\end{enumerate}

%	\begin{enumerate}
%		\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
%		\rangle. $
%		\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
%		\item We implement unitary $V_0 $ or $ V_1$.
%		\item We prepare the measurement $\Delta$ in standard basis (already exists on 
%		Rigetti architecture) on each qubits.
%		\item We calculate the probability of correct discrimination in the following 
%		way
%		\begin{equation}
%		p = ...
%		\end{equation}
%		%\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}
%	\end{enumerate}
\end{scheme}

@@

\subsubsection{Mathematical preliminaries}

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity notation, square matrices will be 
denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$ 
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite 
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is 
positive semidefinite operators having
trace equal to one, will be denoted $\Omega_d$.  The subset of $M_d$ consisting of 
unitary matrices will be denoted by
$\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$.  We will also need a linear
mapping transforming $M_{d_1}$ into $M_{d_2}$, which will be denoted \begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation}

Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
expressed as the action of a quantum channel
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$ Every von Neumann measurement  $\PP_{U}$ can be also rewritten as $\Delta 
\circ \Phi_{(UE)^\dagger}$, where $\Delta$ denotes the completely dephasing 
channel.

\subsubsection{Discrimination of a quantum operations}
This section defines a norm on the space of linear mappings, known as the diamond norm,
and establishes some of its properties. 
Next, we will also explain the precise connection between this
norm and the task of quantum operations discrimination.

Let us now consider a  linear mapping transforming  square matrices into square 
matrices 
that is $\Phi: M_{d_1} \to M_{d_2}$. 
 We define the diamond norm of a map $\Phi$ as 
\begin{equation}
\|\Phi\|_\diamond = \max_{\|X\|_1=1} \| \left(\Phi \otimes \1\right) (X) \|_1.
\end{equation}

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum operations, $\TT$  and $\SS$, 
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{succ} =  \frac12 + \frac14 \| \TT - \SS \|_\diamond.
\end{equation}
